In order to motivate the Bayesian model that we will develop, we begin with the
premise, also shared by virtually all of the studies mentioned above, that the
memory representation of a text is a representation of its semantic content.  We
assume that, to a first approximation, the text's semantic content can be
described by the set of *discourse topics* that are instantiated in the text,
with each discourse topic being represented by a cluster of inter-related
words. As an example, consider the following text:
  
  > *Despite the challenges associated with building up large-scale production, the*
  > *Japanese are forging ahead in display technology — and for very good reasons.*
  > *Obviously the straightforward replacement of bulky cathode ray tubes with flat*
  > *panel display leads to significant power and weight reductions as exemplified in*
  > *the commercialisation of laptop computers...*
  
\noindent It is clear that this text is about, roughly speaking, topics such as
*industry*, *technology*, and so on.  Likewise, a topic such as *industry*
could in general be exemplified by words like *plant*, *company*, *factory*,
*production*, and many others, while a topic such as *technology* could in
general be exemplified by words like *computer*, *electronics*, *engineering*,
and many others.  In other words, we can approximately describe the semantic content of any given
text by identifying
within it a set of discourse topics, taken from a potentially large set of
discourse topics, with each of these discourse topics being characterized by a
possibly large set of semantically inter-related words. Whenever we read or hear a text, we infer its discourse
topics by recognizing different clusters of words, each of which signify
different topics. These inferred topics then describe our representation of
what the text is about.  On the basis of this representation, recall or
recognition of items from the text is then essentially assessing whether any
given item is compatible with, or typical of, the topics that have been
inferred in the text.  For example, on the basis of the text above, there might
be high probability of recognizing or recalling words such as *display* (which
did occur) or *monitor* (which did not occur) because both are typical of some
of the main inferred discourse topics.

To put the foregoing description in somewhat more formal terms, we will denote
the set of all possible discourse topics across spoken and written language by
$\phi_1, \phi_2 \ldots \phi_k \ldots$, with each $\phi_k$ being,
in general, a probability distribution over all words in the language. Given
that these topics are essentially statistical patterns that characterize spoken
and written language, they can therefore be inferred (or learned, acquired, etc.) 
from a language corpus via Bayes's rule. In other words, if $\data$ is a language corpus, 
using Bayes's rule, as we will explain,  we may infer 
$$
\Prob{\phi_1, \phi_2 \ldots \phi_k \ldots \given \data},
$$
which gives the probability distribution over the topics that characterize $\data$.
Having learned the topics that characterise the language, 
for any particular text, denoted by $\textj{j}$, we can then infer, 
again via Bayes's rule,
$$
\Prob{\pi_j \given \textj{j}},
$$
where $\pi_j$ is a probability distribution over the 
discourse topics in the language. 
This is equivalent to performing a
statistical parse of the topics in the text, and leads to a representation of
what the text is about in terms of a probability distribution over all possible
topics. 
Effectively, therefore, $\pi_j$ is the representation 
of $\textj{j}$ in terms of the topics that characterize or typify it.
Finally, on the basis of this memory representation, using Bayesian posterior
predictive inference, we may predict which words are and are not typical or
compatible with the text's inferred topics, and from this we can calculate the
probability that any given word will, whether falsely or veridically, be recognized
or recalled from the text.  Specifically, having inferred $\pi_j$, we may then calculate
$$
\Prob{w \given \textj{j}} = 
\int\! \Prob{w \given \pi_j} \Prob{\pi_j \given \textj{j}}\ d\pi_j,
$$
which gives the probability distribution over the words that 
are predicted or implied by the $\pi_j$ representation of text $\textj{j}$.

This description provides the outline of a computational version of the general
account of text memory mentioned above. We provide a schematic of this in
Figure \ref{fig:schematic}b.  It also outlines formal or computational
descriptions of the various details that remained vaguely described in the
original account.  For example, according to this computational account, the
necessary background knowledge for text memory is knowledge of statistical
patterns in spoken and written language. We acquire knowledge of these patterns
via Bayesian learning from the statistics of natural language. Our cognitive or
memory representation of any given text is a representation of which topics are
inherent in that text, and this is inferred using posterior inference using
Bayes's rule. Finally, recognizing or recalling items from a text is based on
posterior predictive inference of which words are or are not typical of the
topics inferred in the text.

## A Hierarchical Dirichlet Process Topic Model

The Bayesian model described thus far assumes a probabilistic language model. In
other words, it assumes that we have explicitly specified a probabilistic model
that generates a language corpus, and one that does so specifically in terms of
a set of component distributions $\Phi = \phi_1, \phi_2 \ldots \phi_k \ldots$,
each of which is simply a probability distribution over a vocabulary of words.
This language model, by definition, effectively provides a complete account of
the probabilistic or statistical structure of language.  There are infinitely
many probabilistic language models that we could employ here.  Some, such as
those involving probabilistic grammars, focus on providing accurate descriptions
of more fine-grained syntactic structures. Other probabilistic language models,
such as *probabilistic topic models* [see, for example,
@griffiths:psychrev;@steyvers2007probabilistic;@blei2012probabilistic;@blei03]
have proved highly effective in capturing the statistical patterns that
characterize the coarse-grained statistical structures across spoken and written
language. Here, we use a type of probabilistic topic model known as a
*hierarchical Dirichlet process topic model* (\hdptm)
[@TehJorBea2004b;@teh:JASA].

A \hdptm is a *bag of words* probabilistic language model.  According to a
bag of words model, a language corpus $\data$ is a set of texts, where each text is an
unordered set of words. In other words, within each text, all sequential or
syntactic information is ignored and the only information that is used is which
words occur and with what frequency. More formally, according to a bag of words
model, the corpus of language data $\data$ is treated as a set of $J$ texts $\textj{1},
\textj{2} \ldots \textj{j} \ldots \textj{J}$, where each $\textj{j}$ is a set of
$n_j$ words from a finite vocabulary, represented simply by the $V$ integers
$\{1, 2 \ldots V\}$. In other words, each $\textj{j} =
w_{j1}, w_{j2} \ldots w_{ji} \ldots w_{jn_j}$, with each $w_{ji} \in \{1 \ldots
V\}$, and so $\data = \{w_{j1}, w_{j2} \ldots w_{ji} \ldots w_{jn_j}\}_{j=1}^J$.

As a generative model of this corpus, the \hdptm treats each observed word
$w_{ji}$ as a sample from one of infinitely many discourse topics $\phi_1,
\phi_2 \ldots \phi_k \ldots$, where each $\phi_k$ is a probability distribution
over $\{1 \ldots V\}$. Although each $\phi_k$ is a distribution over the entire
vocabulary, because it will place most of its probability mass on a relatively small
number of words, in effect it identifies a cluster of inter-related words that describe some 
meaningful discourse topic (see examples of inferred topics on Page \pageref{tab:topics}).
The identity of the particular topic distribution from which each $w_{ji}$ is
drawn is determined by the value of a discrete latent variable $x_{ji} \in \{1,
2 \ldots k \ldots\}$ that corresponds to $w_{ji}$. As such, we model each
$w_{ji}$ as $$w_{ji}\vert x_{ji}, \phi \sim \dcat{\phi_{[x_{ji}]}},$$ where
$\dcat{\cdot}$ denotes a categorical distribution. The probability distribution
over these possible values of $x_{ji}$ is given by by $\pi_j$, which is a
categorical distribution over the positive integers that is specific to text
$\textj{j}$. In other words, $$x_{ji} \given \pi_j \sim \dcat{\pi_j}.$$ Each
$\pi_j$, in turn, is assumed to be drawn from a Dirichlet process prior [see @ferguson1973bayesian]
whose
base distribution, $m$, is a categorical distribution over the positive integers
and whose scalar concentration parameter is $a$: $$\pi_j\vert a,m \sim
\ddp{a,m},$$ which is equivalent to
$$\pi_{j1},\pi_{j2}\ldots\pi_{K},\textstyle\sum_{k > K} \pi_k\vert a,m \sim \ddirich{am_1,am_2\ldots am_K, a\textstyle\sum_{k > K} m_k },$$
where $\ddirich{\cdot}$ is a standard Dirichlet distribution, which is a
probability distribution over finite categorical distributions. In other words,
the Dirichlet process prior over each $\pi_j$ is the infinite dimensional
counterpart to a Dirichlet distribution over finite dimensional probability
distributions. The $m$ base distribution of the Dirichlet process is assumed to
be drawn from a stick-breaking prior [see @ishwaran2001gibbs] with a parameter
$\gamma$: $$m\vert\gamma \sim \dstick{\gamma}.$$ A stick-breaking distribution
is a probability distribution over infinite-dimensional categorical
distributions, and is the infinite dimensional counterpart of the Generalized
Dirichlet Distribution [@connor1969concepts]. The prior distributions of the
Dirichlet process concentration parameter $a$ and the stick-breaking parameter
$\gamma$ are Gamma distributions, both with shape and scale parameters equal to
1. For the topic distributions, $\phi_1, \phi_2 \ldots \phi_k \ldots$, we can
assume they are independently and identically drawn from a Dirichlet
distribution with a length $V$ location parameter $\psi$ and concentration
parameter $b$. In turn, $\psi$ is drawn from a symmetric Dirichlet distribution
with concentration parameter $c$ 
(i.e. $\psi\vert c \sim \ddirich{c\cdot \mathbf{1}_V}$, 
where $c \cdot \mathbf{1}_V$ is a length $V$ vector with each element being $c$). 
Finally, both $b$ and $c$, like $a$ and
$\gamma$, can be given Gamma priors, again with shape and scale distributions
equal to 1. The full specification of the \hdptm is provided in Figure \ref{fig:generative_model}.
```{r schematic, child = 'generativemodel.Rmd'}
```



### Inference

Given any corpus of natural language $\data$, 
the posterior distribution over all the unobserved variables, which includes 
all latent variables, parameters, and hyperparameters, 
in the \hdptm is
$$
\begin{split}
&\Prob{\vec{x}_{1:J}, \phi, \pi_{1:J}, \psi,  m, a, b, c, \gamma \given \data}\\
&\phantom{\propto}\propto 
\Prob{\psi\given c}\Prob{\pi_{1:J}\given a,m}\Prob{m\given\gamma}\Prob{b,c,a,m}
\int\!\Prob{\data \given \vec{x}_{1:J}, \phi}\Prob{\vec{x}_{1:J} \given \pi_{1:J}}\Prob{\phi\given b, \psi} \ d\vec{x}_{1:J}d\phi
\end{split}
$$
where $\vec{x}_{1:J} \doteq \{x_{j1}, x_{j2} \ldots x_{jn_j}\}^{J}_{j=1}$, 
$\phi \doteq \phi_1, \phi_2 \ldots$, 
$\pi_{1:J} \doteq \pi_1, \pi_2 \ldots \pi_j \ldots \pi_J$. We
may draw samples from this high dimensional posterior distribution using a Gibbs
sampler. We derived this sampler based on the work of @TehJorBea2004b, @teh:JASA,
@newman2009distributed and we describe it in full technical detail in
@andrews_2019.


### Posterior predictive distributions

Having inferred the posterior distribution of the unobserved variables in the \hdptm, 
for any new text $\textj{{j^\prime}}$, 
the posterior predictive distribution over the vocabulary conditioned on $\textj{{j^\prime}}$ is,
informally speaking, the distribution over words that are consistent with the discourse
topics of text $\textj{{j^\prime}}$. It is calculated as follows:
$$
\begin{aligned}
\mathrm{P}(w \given \phi, \textj{{j^\prime}}, a, m) &= \int \mathrm{P}(w \given \phi, \pi_{j^\prime}) \mathrm{P}(\pi_{j^\prime} \given \textj{{j^\prime}}, \phi, a, m) d\pi_{j^\prime},\\
&= \int \bigg[ \sum_{k=1} \mathrm{P}(w \given \phi, x=k)\mathrm{P}(x=k \given \pi_{j^\prime}) \bigg] \mathrm{P}(\pi_{j^\prime} \given \textj{{j^\prime}}, \phi, a, m) d\pi_{j^\prime}
\end{aligned}
$$
where $\mathrm{P}(\pi_{j^\prime} \given \textj{{j^\prime}}, \phi, a, m)$ is the posterior
distribution over topic distributions for text $\textj{{j^\prime}}$. 
We can sample from the posterior $\Prob{\pi_{j^\prime} \given \textj{{j^\prime}}, \phi, a, m}$ 
using a blocked Gibbs sampler that iterates the following two sampling steps:
$$
\begin{aligned}
\vec{x}_{j^\prime} &\sim \Prob{\vec{x}_{j^\prime} \given \vec{w}_{j^\prime},\phi, \pi_{j^\prime}},\\
\pi_{j^\prime} &\sim \Prob{\pi_{j^\prime} \given \vec{x}_{j^\prime}, a, m},
\end{aligned}
$$
where $\vec{w}_{j^\prime} = w_{j^\prime 1}, w_{j^\prime 2} \ldots w_{j^\prime
n_{j^\prime}} = \textj{{j^\prime}}$ and $\vec{x}_{j^\prime} = x_{j^\prime 1},
x_{j^\prime 2} \ldots x_{j^\prime n_{j^\prime}}$. In the first step, we sample
from the posterior distribution over the latent variables that correspond to the
words in text $\textj{{j^\prime}}$. In the second step, we sample from the
posterior distribution over $\pi_{j^\prime}$ given the values of
$\vec{x}_{j^\prime}$ sampled in the previous step. The posterior distributions
in both of these two steps can be calculated exactly. 


## Implementation and Simulations

In order to explore and evaluate the model just described, we implemented it as
a Python/Fortran program [@gustavproject:19.4.15] and trained it on a large
representative corpus of natural language. The corpus was obtained by extracting
paragraphs, or concatenations of consecutive paragraphs, of between 250 and 500
words from the British National Corpus (\bnc) [@bnc]. After removing any texts
that contained as sub-texts those texts that were used as the stimuli in the
behavioural experiment (see next section), this resulted in a set of 183,975
texts of between 250 and 500 words each, with a collective total of over 78
million word tokens. After removing stop words, words that occurred less than 5
times in total in the, and words that did not occur in a standard dictionary of
English words[^standard_dictionary], there was a total vocabulary of 49,328 word types in the corpus.
See Appendix A for details on the Python code used to create the corpus, and for
how to obtain a copy of the corpus itself.

[^standard_dictionary]: We used the `2of4brif.txt` word list 
from the `12Dicts` collection of English word lists, available 
at `http://wordlist.aspell.net/12dicts`.\label{footnote_standard}

We trained the \hdptm using the corpus just described using the Gibbs sampler.
This drew samples from posterior distribution over $\vec{x}_{1:J}, \phi,
\pi_{1:J}, \psi,  m, a, b, c, \gamma$, for 20,000 iterations. After approximately 
a few hundred iterations, 99\% of the posterior probability mass converged on
approximately 1,340 topics. In other words, after a few hundred
iterations
$$
\sum_{k=1}^{K=1340} m_{s_i} \approx 0.99,
$$
where $s_1, s_2 \ldots$ are the indices of the elements of $m$ sorted in
decreasing order. Samples from iteration 19,000 to 20,000, thinning by a factor
of 10, were then retained as the final set of posterior samples.
Using these posterior samples, for each of held-out text $\textj{{j^\prime}}$, we 
calculated the posterior predictive distributions as follows:
$$
\mathrm{P}(w \given \bar{\phi}, \textj{{j^\prime}}, \bar{a}, \bar{m}),
$$
where $\bar{\phi}$, $\bar{a}$, $\bar{m}$ are the means of the posterior samples of 
$\phi$, $a$, $m$, respectively. 

## How the Bayesian Model of Text Memory Works

To reiterate, the contention of this paper is that the widely held, yet arguably
vague and untestable, account of memory for text (see Figure
\ref{fig:schematic}a) can be described formally as a Bayesian computational
model (see Figure \ref{fig:schematic}b), and in practice we can instantiate this
model using a \hdptm. Having described the \hdptm, we may now describe how it
can act as a computational model of human memory for text. There are three main
components to the model. First, there is the representation of the background
knowledge that is relevant for constructing memory representations. Second,
there is how this background knowledge is used, via Bayes's rule, to infer a
representation of a text's semantic content. This is effectively the text's
memory representation. Finally, having inferred the text's representation, the
recognition or recall of the word's in the text is based on the posterior
predictive distribution over words conditional of the text's represenation.

The model assumes that background knowledge, or at least the background
knowledge that is most relevant for inferring a text's semantic content, is
represented by a large set of discourse topics $\phi_1, \phi_2 \ldots$. These
are assumed to be learned over the course of our lives from experience with the
statistics of spoken and written language, and are the Bayesian model's
counterparts of the *schemata* or *scripts*, etc., in the widely held account of
text memory. Examples of topics inferred by the \hdptm trained on the corpus
described above are illustrated in Table \ref{tab:topics}. In each case, we
display the 10 most probable words[^foot_displaying_topics] in each topic. In
general in a \hdptm or any other probabilistic topic models, each topic often
identifies some intuitively meaningful discourse topic that could perhaps be
provided with a meaningful approximate label such as, for some of the examples
shown, *university*, *home furniture*, *Northern Ireland*, and so on. Although
there are in principle infinitely many topics in a \hdptm, in any given model,
the posterior distribution will be concentrated on a relatively small number of
topics. For any medium sized corpus of natural language, we would expect the
posterior distribution to be concentrated on at least many hundreds, though
possibly thousands, of topics like those shown in the table. In the simulations
that we use in this paper, as mentioned above, approximately 99% of the
posterior distribution over the topics is concentrated on around 1340 distinct
topics.

```{r, results='asis'}

x_table <- read_csv('example_topics.txt', col_names = F) %>% 
  as.matrix() %>% 
  t() %>% 
  .[, sample(ncol(.))] %>% 
  matrix(nrow = 5 * 10, ncol = 7) %>% 
  as.data.frame(row.names=F) %>% 
  xtable(align='cccccccc',
         label='tab:topics',
         caption = 'Examples of topics inferred by the \\hdptm trained on the \\bnc.
         In each case, the 10 most probable words in the topic are shown.
         ')

print(x_table, 
      include.rownames=F,
      include.colnames=F,
      size="\\footnotesize\\selectfont",
      floating = TRUE, 
      table.placement = NULL,
      latex.environments = "center",
      hline.after = c(0,10,20,30,40,nrow(x_table)))




```

Having learned the repertoire of discourse topics, we can infer the probability distribution over the topics in any new text $\textj{{j^\prime}}$ using Bayes's rule: $\Prob{\pi_{j^\prime} \given \textj{{j^\prime}}, \phi, a, m}$.
Here, $\pi_{j^\prime}$ is a probability distribution over all topics and can be
seen as the semantic representation of text $\textj{{j^\prime}}$: it tells us
which topics are instantiated in text $\textj{{j^\prime}}$ and with what
probabilities. In other words, it provides a semantic parse of text
$\textj{{j^\prime}}$ in terms of the repertoire of discourse topics. As an
example, consider the following text.

> *Individual goals may well be in conflict with organisational goals even where*
> *there is a strong commitment to one goal by the organisation. It is worth*
> *repeating that there may also be competing goals within the organisation itself.*
> *Thus, for example, the prison service may wish to pursue the twin goals of the*
> *punishment of offenders  and at the same time their rehabilitation. These two*
> *goals may prove to be incompatible and yet each will have its supporters. Within*
> *the National Health Service there may well be a commitment to provide the best*
> *possible medical care that the state of medical knowledge will allow and yet*
> *this may be tempered by the requirement to act within tight financial*
> *constraints. The two goals may be in conflict. Many organisations will face the*
> *problem of which goals to prioritise in the light of competing goals.*

\noindent Using the averages of the samples of the posterior of $\phi$, $a$,
$m$, which we denote by $\bar{\phi}$, $\bar{a}$, and $\bar{m}$, respectively,
we can then draw samples from $\Prob{\pi_{j^\prime} \given \textj{{j^\prime}},
\bar{\phi}, \bar{a}, \bar{m}}$, the average of which we'll denote by
$\bar{\pi}_{j^\prime}$. In the following table, we show the 7 most probable
topics and their probabilities. 

\setlength{\parindent}{0pt}

```{r, results='asis'}
x_table <- read_csv('text_24_vpi.txt', col_names = F) %>% 
  as.matrix() %>% 
  t() %>% 
  as.data.frame(row.names=F) %>% 
  xtable(align='cccccccc')

cat('\\begin{center}',
print(x_table, 
      include.rownames=F,
      include.colnames=F,
      comment = F,
      print.results = F,
      size="\\footnotesize\\selectfont\\noindent",
      floating = FALSE, 
      table.placement = NULL,
      latex.environments = "center",
      hline.after = c(0,1,nrow(x_table))
),
'\\end{center}')

```
\noindent This shows that this text is represented as being about a mixture of topics,
including those that might be labelled *objectives* (with probability 0.31), *conflict* (0.12),
*organization* (0.12), and so on. 

\setlength{\parindent}{30pt}

Finally, having inferred the distribution over the topics of text
$\textj{{j^\prime}}$, the recognition or recall of words from this text is based
on the posterior predictive distribution over words conditional on
$\pi_{j^\prime}$. If $\tilde{\pi}^1_{j^\prime}, \tilde{\pi}^2_{j^\prime},
\ldots \tilde{\pi}^t_{j^\prime} \ldots \tilde{\pi}^t_{j^\prime}$ are $T$ samples from
the posterior distribution $\Prob{\pi_{j^\prime} \given \textj{{j^\prime}}, \bar{\phi}, \bar{a}, \bar{m}}$, then the posterior predictive distribution over the vocabularly is given by the following.
$$
\begin{aligned}
\mathrm{P}(w \given \bar{\phi}, \textj{{j^\prime}}, \bar{a}, \bar{m}) 
&\approx
\frac{1}{T}\sum_{t=1}^T \mathrm{P}(w \given \bar{\phi}, \tilde{\pi}^t_{j^\prime}),\\
&= \frac{1}{T}\sum_{t=1}^T \bigg[\sum_{k=1} \mathrm{P}(w \given \bar{\phi}, x=k)\mathrm{P}(x=k \given \tilde{\pi}^t_{j^\prime})\bigg]
\end{aligned}
$$
\noindent Illustrations of posterior predictive distributions for some example
held out texts, which we then use in the behavioural experiments described next,
are shown in Figure \ref{fig:wordclouds}. In each case, the posterior predictive
distribution is a distribution over the entire vocabulary of words, and tells us
how much any given word is compatible with, or typical of, the topic-based
representation of the text. From this distribution, we have a model of what
words will and will not be remembered, whether veridically or falsely, from any given text.


```{r introduction, child = 'wordcloud.Rmd'}
```




<!-- [^prob-of-vpi]: Given that $\pi_{j^\prime}$ is a probability distribution over all topics, -->
<!-- the posterior distribution is a probability distribution over $\pi_{j^\prime}$. -->

[^foot_displaying_topics]: While each topic is always defined by a probability
distribution over the entire vocabulary of $V$ words, in most topics the
probability mass is concentrated on a relatively small subset of words, and this
allows us to illustrate topics by displaying only their most probable words.



